### 2022-03-24 example latex paper/main.tex
Added example latex file

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Evolving robots to play capture the flag}
\author{Uriel Mandujano and Daniel Redelmeier}
\date{} %don't display the current date

\begin{document}

\maketitle

\begin{abstract}

  In nearly all videogames, designing smart and complex artificial agents is necessary to ensure a fun player experience. In this experiment, we attempt to meet this challenge by training robots to develop effective strategies in a simple two-player game similar to Capture the Flag. The robots start on opposite ends of a field, and must attempt to reach a flag in the center and return it to their base without being touched. The catch is that any robot carrying the flag can only move at 70\% speed, enabling the flag to be stolen. Stealing the flag from the opponent grants a significant advantage: the flag is transferred and the previous flag-carrier is teleported back to their base. We implemented the game in the Pyrobot simulator, and used separate feedforward neural networks to control two robots playing against each other. Using this model, we evolved the robots to continuously improve on their performance through competitive coevolution in NEAT. We hypothesized several strategies that might develop. A straight offensive strategy would involve grabbing the flag as quickly as possible, and trying to bring it back to the base before being caught. A more defensive strategy might make use of the captuing mechanism to steal the flag after the opponent had grabbed it. Our results showed that both strategies evolved at different times (first offensive, and later defensive), as well as others. Although the final best strategies were straightforward and predictable from a gameplay perspective, we believe our setup has strong potential; further experimentation and refinement could produce robots that are worthy of human opponents. Moreover, this experiment shows that competitive coevolution can be used to improve videogame AI as a whole with relatively minimal human input.

\end{abstract}

\section{Introduction}

  Capture the Flag is a children's game usually played outside. A large playing area is divided in two, with each region being a safe area for one team but not the other. Both teams hide an object -- the ``flag" -- somewhere in their region before the game begins. The object of the game is to find the opponent's flag and bring it back to the safe zone before the other team. If an individual is tagged outside of their zone while looking for or carrying the flag, they must return to their side of the field and drop the flag if they have it. Some online versions of Capture the Flag exist, such as TagPro (see \url{tagpro.koalabeast.com}).

  In our simplified version of the game, we ran the same competitive coevolution algorithm on two populations of robot brains to improve overall performance. Due to the competitive nature of the game, a robot tended to have a higher fitness if it was closer to winning the game. The population was capped, so phenotypes of fitter individuals would be more present in the next generation. When evaluating one robot, the brain of its competitor was set as the best-performing opponent phenotype from the previous generation, which ensured that both robots had to develop new strategies to succeed. We performed competitive coevolution using Neuroevolution of Augmenting Topologies (NEAT), ``a method for evolving artificial neural networks with a genetic algorithm," developed by Kenneth O. Stanley in 2002 \cite{NEAT_website}.

We are not the first to run such experiments.  Stanley and Miikulainen ran competitive coevolution tests to create a ``Robot Duel", a simulated conflict between two distinctly evolving populations \cite{NEAT}. Pitting two individuals against each other addresses the issue of getting stuck in local optima that can occur in standard back-propagation for neural networks.  As soon as one robot has evolved a sufficiently dominant strategy, the other robot works to evolve a more dominant strategy, creating a cycle that could continue indefinitely. Another distinct advantage to using a co-evolution strategy is it provides a basis for studying complexification. The network begins with a minimal number of nodes and adds more nodes and connections as strategies are created. In their Robot Duel, they expected to never encounter an entirely dominant strategy simply because evolution is continuous.  Their results comfirmed that continuous evolution results in increasingly sophisticated strategies.

Another study investigating coevolution was done by Floreano and Nolfi \cite{red_queen}. They use pursuit-evasion goals to explore the dynamics of learning. Pursuit-evasion is a competition that models natural occurences, e.g. a cat and mouse.  In our own tests, we experimented with two different fitness functions: one that only offered a positive reward upon the accomplishment of tasks; the other subtracted fitness from an individual if it lost an encounter with the opponent.  The latter fitness function is more akin to Floreano and Nolfi's study and we were curious to see how impactful the change would be. They never observed the absolute dominance of one strategy. In fact, some individuals of `more evolved' generations would lose to those of earlier, `less evolved' generations.  Granted, they compared the most recent generation's individuals to those in simulations ten generations earlier.  But even with this caveat, no strategies  that developed in the most recent simulations exhibited more complex solutions.

Miller and Cliff also experimented with coevolution and pursuit-evasion \cite{pursuit}. They note that if in a simulation, the individuals benefit from being together, they would evolve in a manner more likely to pursue and keep track of each other.  It they benefitted from being seperate, they would evolve to keep their distance at all times.  If, however, one individual benefitted from an encounter with the other, some sort of cat and mouse game would develop. Indeed, this is similar to Floreano and Nolfi's study, and is what we were attempting to replicate in our experiments with the detrimental fitness function.

\section{Experiments}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.5]{Screenshot_-_030714_-_00_15_04.png}
\end{center}
\caption{This picture shows the Pyrobot playing field after a sample trial has just begun. Both robots have already begun to leave their respective bases and move toward the flag, coded as a light source in Pyrobot. The robots have access to their front sonar and light-detection sensor values, as well as 8 other inputs.}
\label{setup}
\end{figure}

Our experiment setup consisted of a 10-by-10 playing field modelled in the Pyrobot simulator. A red and and a blue robot start on opposite sides of the field equidistant from the flag in the center, just in front of their bases (the blue robot had the white base, and the red robot had the black base). The rules of the game were simple:
\begin{itemize}
  \item Both robots could move anywhere on the field, although both bases were impassable
  \item If the robot touched the flag, it immediately gained possession while its maximum speed was reduced by 30\%
  \item If one robot came within 0.75 units of the flag-carrying robot, the flag immediately changed possession, and the initial-flag carrier was teleported back to their starting position
  \item Most trials ran for 125 steps. However, if either robot returned the flag to their base, the trial ended immediately
\end{itemize}

The robots had 11 total input values and three hidden nodes to determine motor output: the front right and left light sensor input, the front center sonar sensor input, two booleans that indicated whether the blue or red robot had the flag, the robot's x and y distance from its base and its heading, and  its x and y distance from the opponent as well as the opponent's heading.

Our fitness function to evaluate a robot's performance was more complex, and gave greater rewards scaling with how close a robot was to achieving the main objective. Initially, the fitness was determined by how close a robot's final position was to the flag, if neither competitor retrieved it. When the flag began to be picked up, different fitnesses were calculated to encourage robots to return the flag if they had it, and pursue their opponent if not. Due to the layered quality of our function, we coded it using several conditional statments:

\begin{verbatim}
  if the robot being evaluated does not win the game in 125 time steps:

     if neither robot ever grabs the flag:
        fitness = 8 - distance from flag
     if the robot never had the flag, but its opponent did:
        fitness = 16 - distance from opponent   
     if the robot had the flag, but its opponent has it at the end of the trial:
        fitness = 20 - distance from opponent  
     if the robot has the flag at the end of the trial:
        fitness = 24 - distance from home base

  else:
     fitness = 24
\end{verbatim}

Each robot was evaluated twice, so the maximum possible fitness (the sum of the fitness for each evaluation) was 48. Using this setup, we ran three similar experiments over three runs each. All runs were done with a population of 40 for each robot over 15 generations, which we found to be sufficient for obtaining useful results. Our first test used the base code and the default evolution parameters. The second test altered the fitness function to punish a robot with -6 fitness if it ever lost the flag. Our final test used the initial fitness function, but increased the probability of adding connections and nodes from 5\% to 10\% and 3\% to 6\% respectively.

\section{Results}

\begin{figure}[h]
\begin{center}
\fbox{\includegraphics[scale=0.3]{Screenshot_-_030614_-_11_10_17.png}}
\fbox{\includegraphics[scale=0.315]{Screenshot_-_030614_-_11_10_46.png}}
\end{center}
\caption{Both graphs show the average and best fitness progression for trial 2 of the first experiment. The average fitness (in blue) is shown for the red robot on the left and the blue robot on the right, and there is a clear upward trend in performance for both. We discovered that the sharp fall in the blue robot's fitness at generation 13 was likely caused by a mostly defensive blue population facing off against a red phenotype that likewise played defensively. When we evaluated that generation specifically, we found that both robots were reluctant to grab the flag, which explains the decline in fitness. This was remedied in the subsequent generation, as a more offensive blue population had evolved.}
\label{experiment1fitness}
\end{figure}

\begin{figure}[h]
\begin{center}
\fbox{\includegraphics[scale=0.295]{40.png}}
\fbox{\includegraphics[scale=0.32]{34.png}}
\fbox{\includegraphics[scale=0.26]{cooler_network.png}}
\end{center}
\caption{Like Figure 2, the graphs show the average and best fitness progression for the red (top-left) and blue (top-right) robts. These graphs came from trial 3 of the second experiment. As expected, there is still a net increase in fitness for each robot. Interestingly, the blue robot developed a unique strategy that we did not expect or find in other trials. Beginning in Generation 9, the blue robot would circle to the left of the red robot instead of heading straight for the flag. When the red robot grabbed the flag and began to rotate back to its base, the blue robot was already in a position to intercept. This is where the robot's strategy began to disintegrate, although this may have been remedied with more generations. The most-fit blue topology from Generation 12 is shown below the graphs. Note the addition of a hidden node (17) that was not present at the beginning.}
\label{experiment1fitness}
\end{figure}

Our results for all three experiments showed general increases in fitness for both the red and blue robots as evolution progressed. Specific evolutionary developments are shown in Figures 2 and 3 for Experiments 1 and 2. Unfortunately, none of the trials in Experiment 3 had particularly interesting results, and the average fitness increase was was lowest of all the experiments. This might have been different if we ran our trials for more generations, as the additional nodes and connections might have had a more positive impact. This is quite plausible, since complexifying coevolution in NEAT has already been shown to be effective.

\section{Discussion}

Our results add further evidence supporting the effectiveness of using competitive coevolution to create solutions to problems in AI. We consistently saw the fitness improve thoughout each trial.  For example, early successful phenotypes tended to compete in a mad dash to the flag in the center.  Later on, we saw both robots become more patient and hesitate before going for the flag.  Another strategy was to wait for the opponent to obtain the flag before "capturing the flag" and resetting their position. Further testing with longer trial runs may yield more interesting and unforseen strategies and the development of complex strategies falls in line with the current research \cite{NEAT}.  Initially the robots develop simple strategies corresponding to few nodes and connections. Building off of these networks creates foundations for more creatively complex strategies.

Overall, we believe that our experiments were successful in developing adaptive strategies without the need to hard-code functions to determine motor output. If we are able to run more generations and develop better topologies, we may see the rise of robots that can reasonably compete against a human-controlled opponent. We could then evaluate and evolve the robots based on interactions with humans, and reveal insights into human tendencies for strategies in the game. We believe that competitive coevolution should be applied to other challenged in AI development outside of the Pyrobot simulation.

\section{Acknowledgements}

Our experiment was more complicated than initially intended, and we could not have done it by ourselves. We would like to thank Lisa Meeden for providing the base code to perform competitive coevolution in NEAT on a Pyrobot simulation, as well as guidance on the project. Additionally, we would  like to thank Joe Boninger for showing us a trigonometric algorithm to help the robots make better turning decisions. Finally, thanks to Ben Xie and Tyler Zon for helping us create separate files that ran our tests outside of the simulation window, which allowed us to perform more trials in a shorter amount of time.

\begin{thebibliography}{1}

  \bibitem{red_queen} Dario Floreano and Stefano Nolfi, God save the red queen! competition in co-evolutionary robotics. {\em Evolutionary Computation}, 5, 1997.

  \bibitem{pursuit}  Geoffrey F. Miller and Dave Cliff, Co-evolution and pursuit and evasion i: Biological and game-theoretic foundations. 1994.

  \bibitem{NEAT} Kenneth O. Stanley and Risto Miikkulainen, Competitve coevolution through evolutionary complexification {\em HJournal of Artifical Intlelignce Research}, 21, 2004.

  \bibitem{NEAT_website} Kenneth O. Stanley. The NeuroEvolution of Augmenting Topologies (NEAT) Users Page. {\em NEAT Software FAQ}, 2013. \url{http://www.cs.ucf.edu/~kstanley/neat.html}

  \end{thebibliography}

\end{document}

### 2022-03-25 paper.tex
Base template of pdf

\documentclass{article}

\usepackage{listings}

\begin{document}
\author{Dalton Kajander, Shawn Guydeene, Blake Lastname}
\title{Concurrent Hash Tables in Java}
\date{\today{}}

\maketitle

\begin{lstlisting}[language={Java},caption=Testing surce code in \LaTeX{}.,breaklines=true,frame=single]
class HashTableCon {
    public static void main(String[] args) {
        bool isConcurrent = true;
    }
}
\end{lstlisting}

\end{document}

### 2022-03-25 paper.tex
Ignore .toc
Basic sections

@@
 \date{\today{}}
-
 \maketitle
 
+\newpage
+\tableofcontents
+\newpage
+
+\section{The Problem}
+What problem?
+
+\section{Research}
+Oh, that problem.
+
+\section{Related work}
+I've seen a similar problem before
+
+\section{Contributions}
+I've fixed this problem once
+
+\section{Algorithms}
+And this is how I did it!
 \begin{lstlisting}[language={Java},caption=Testing surce code in \LaTeX{}.,breaklines=true,frame=single]
-class HashTableCon {
+class HashTable<T> {
+    bool isConcurrent;
     public static void main(String[] args) {
-        bool isConcurrent = true;
+        isConcurrent = true;
+    }
+
+    public HashTable<T>(bool isConcurrent) {
+        this.isConcurrent = isConcurrent;
     }
@@
 
+\section{Experimental Results}
+Well, it worked on my machine.
+
 \end{document}

### 2022-03-25 paper.tex
Project Description and Requirements page

@@
 \usepackage{listings}
+\usepackage{hyperref}
 
@@
 
+\newpage
+\section{Description and Requirements}
+The purpose of the project is to give you a chance to explore in some depth a topic that is related to the subject matter of this course.
+The project consists of two parts, an activity and a report.
+Your final report must present original work on a challenging problem.
+\begin{enumerate}
+    \item The activity involves scholarly work—exploring a topic, reading relevant literature, writing software, and running experiments—and must include either a substantial amount of programming or a strong theoretical mathematical treaty of issues related to multicore programming.
+    \item The final report is a short paper (6–10 pages) describing the results of the activity and should reflect the substantial amount of work that you put into the activity.
+The paper should make clear what is the problem, the state-of-the art in this research area, a discussion of related work, a clear definition of the contributions of your paper, a detailed discussion of your algorithms and approach and their application, and performed experimental results.
+The report will be graded for writing quality as well as for technical content, so attention should be paid to organization, grammar, spelling, and scholarly style.
+All references used, both text and code, must be properly cited in the bibliography.
+Your are required to use LaTeX (\url{http://en.wikipedia.org/wiki/LaTeX} (Links to an external site.)) to prepare your report.
+    \item Source code: typically in the course of your research you will develop source code for your algorithms and experimental results.
+Please include with your mid-term and final reports the following:
+    \begin{enumerate}
+        \item The source code of your algorithms along with a brief README file
+        \item The source code of your experiments along with a brief README file
+        \item The source code of your paper (in a LaTeX format).
+    \end{enumerate}
+\end{enumerate}
+\textit{Submission} \\
+All materials should be submitted online using Webcourses as a single zip or rar archive. \\
+\textit{Format} \\
+All papers should follow either the ACM or the IEEE Manuscript Format. \\
+A sample IEEE template and further formatting instructions are available here: \\
+\url{http://www.ieee.org/conferences_events/conferences/publishing/templates.html}
+
 \end{document}

### 2022-03-25 paper.tex
Hyphen fix

@@
     \item The activity involves scholarly work—exploring a topic, reading relevant literature, writing software, and running experiments—and must include either a substantial amount of programming or a strong theoretical mathematical treaty of issues related to multicore programming.
-    \item The final report is a short paper (6–10 pages) describing the results of the activity and should reflect the substantial amount of work that you put into the activity.
+    \item The final report is a short paper (6-10 pages) describing the results of the activity and should reflect the substantial amount of work that you put into the activity.
 The paper should make clear what is the problem, the state-of-the art in this research area, a discussion of related work, a clear definition of the contributions of your paper, a detailed discussion of your algorithms and approach and their application, and performed experimental results.

### 2022-03-25 paper.tex
Added Last Name
Conversion to IEEE formatting

@@
-\documentclass{article}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
 
@@
 \usepackage{hyperref}
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\begin{document}
+\title{A Look at Implementing Councurrent Hash Tables in Java}
 
-\begin{document}
-\author{Dalton Kajander, Shawn Guydeene, Blake Lastname}
-\title{Concurrent Hash Tables in Java}
-\date{\today{}}
+\author{\IEEEauthorblockN{ Shawn Guydeene}
+\and
+\IEEEauthorblockN{ Dalton Kajander}
+\and
+\IEEEauthorblockN{ Blake Celestian}
+}
+
 \maketitle
 
-\newpage
-\tableofcontents
-\newpage
+\begin{abstract}
+Abstract goes here.
+\end{abstract}
+
+\begin{IEEEkeywords}
+Concurrent Programming, Hash-table representations
+\end{IEEEkeywords}
 
 \section{The Problem}
-What problem?
+Our group looked into the creation of hash tables in Java, and how one would be able to
 

### 2022-03-25 main/HashTable.java
Create HashTable.java

@@ -1,0 +1,115 @@
+class HashTable {
+    public static final int ARR_SIZE = 128;
+
+    private int[] table = new int[ARR_SIZE];
+    private boolean[] nulls = new boolean[ARR_SIZE];
+    private boolean[] cleans = new boolean[ARR_SIZE];
+    private int capacity;
+
+    public static void main(String[] args) {
+
+    }
+
+    public HashTable() {
+        for (int i = 0; i < ARR_SIZE; i++) {
+            table[i] = 0;
+            nulls[i] = true;
+            cleans[i] = true;
+        }
+        capacity = 0;
+    }
+
+    public void put(int item) {
+        if (capacity == ARR_SIZE)
+            return;
+        int key = hash(item);
+
+        // System.out.println("Item: " + item + ", key: " + key);
+
+        if (nulls[key]) {
+            // System.out.println(key + " is null index (" + item + ")");
+            table[key] = item;
+            nulls[key] = false;
+            cleans[key] = false;
+        } else {
+            // System.out.println(key + " is not null index (" + item + ")");
+            int i = 1;
+            while (!nulls[key + (i * i)]) {
+                // System.out.println("p: " + (key + (i * i)) + " is not null index (" + item +
+                // ")");
+                i++;
+            }
+            // System.out.println("p: " + (key + (i * i)) + " is null index (" + item +
+            // ")");
+            table[key + (i * i)] = item;
+            nulls[key + (i * i)] = false;
+            cleans[key + (i * i)] = false;
+        }
+
+        capacity++;
+    }
+
+    public boolean search(int item) {
+        int key = hash(item);
+        int i = 0;
+
+        while (!cleans[key + (i * i)]) {
+            if (item == table[key + (i * i)] && !nulls[key + (i * i)]) {
+                return true;
+            }
+            i++;
+        }
+
+        return false;
+    }
+
+    public void remove(int item) {
+        int key = hash(item);
+        int i = 0;
+
+        if (!search(item)) {
+            return;
+        }
+
+        while (!cleans[key + (i * i)]) {
+            if (item == table[key + (i * i)]) {
+                nulls[key + (i * i)] = true;
+                table[key + (i * i)] = 0;
+            }
+            i++;
+        }
+        capacity--;
+    }
+
+    public int getSize() {
+        return capacity;
+    }
+
+    public boolean isEmpty() {
+        return this.getSize() == 0;
+    }
+
+    // Need to make a better hash function
[…]

### 2022-03-25 main/HashTable.java
Commented

@@ -1,16 +1,17 @@
 class HashTable {
-    public static final int ARR_SIZE = 128;
+    public static final int ARR_SIZE = 128; // The maximum capacity of a HashTable
 
-    private int[] table = new int[ARR_SIZE];
-    private boolean[] nulls = new boolean[ARR_SIZE];
-    private boolean[] cleans = new boolean[ARR_SIZE];
+    private int[] table = new int[ARR_SIZE]; // Where values are stored
+    private boolean[] nulls = new boolean[ARR_SIZE]; // If a value is null, it is true at the same index here
+    private boolean[] cleans = new boolean[ARR_SIZE]; // Where clean indices are stored
     private int capacity;
 
     public static void main(String[] args) {
-
+        System.out.println("Hello, world!");
     }
 
     public HashTable() {
+        // Generate arrays, each with default values. Set capacity to 0
         for (int i = 0; i < ARR_SIZE; i++) {
             table[i] = 0;
             nulls[i] = true;
@@ -19,40 +20,44 @@ class HashTable {
         capacity = 0;
     }
 
+    // Add an item to the hash table
     public void put(int item) {
-        if (capacity == ARR_SIZE)
+        // If capacity if full, do not add
+        if (capacity >= ARR_SIZE)
             return;
+
+        // Get key
         int key = hash(item);
 
-        // System.out.println("Item: " + item + ", key: " + key);
-
+        // Check if first try addition
         if (nulls[key]) {
-            // System.out.println(key + " is null index (" + item + ")");
             table[key] = item;
             nulls[key] = false;
             cleans[key] = false;
-        } else {
-            // System.out.println(key + " is not null index (" + item + ")");
+        }
+        // Otherwise, quadratically probe for empty indices
+        else {
             int i = 1;
             while (!nulls[key + (i * i)]) {
-                // System.out.println("p: " + (key + (i * i)) + " is not null index (" + item +
-                // ")");
                 i++;
             }
-            // System.out.println("p: " + (key + (i * i)) + " is null index (" + item +
-            // ")");
             table[key + (i * i)] = item;
             nulls[key + (i * i)] = false;
             cleans[key + (i * i)] = false;
         }
 
+        // Increase capacity
         capacity++;
     }
 
+    // Search the hash table for a specific item, return true if found, else false
     public boolean search(int item) {
         int key = hash(item);
         int i = 0;
 
+        // Quadratically probe for the item
+        // Dirty indices are considered to be a collision, the item may of been probed
[…]

### 2022-03-25 paper.tex
Finished Introduction

@@
 \usepackage{hyperref}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \begin{document}
-\title{A Look at Implementing Councurrent Hash Tables in Java}
+\title{An Implementation of Councurrent Hash Tables in Java}
 
@@
 
-\section{The Problem}
-Our group looked into the creation of hash tables in Java, and how one would be able to
+\section{Introduction}
+The problem that our group chose to tackle, is how one would parallelize a hash table, and see how much of a performance increase,
+if any would arise from such an implementation. This paper seeks to explain the how hash tables function as a data structure, previous
+attempts at parallelizing them, our group's attempt at creating a concurrent hash table, and experimental data collected showing the
+effects of parallelizing on runtime for the program.
 
@@
 
-\newpage
-\section{Description and Requirements}
-The purpose of the project is to give you a chance to explore in some depth a topic that is related to the subject matter of this course.
-The project consists of two parts, an activity and a report.
-Your final report must present original work on a challenging problem.
-\begin{enumerate}
-    \item The activity involves scholarly work—exploring a topic, reading relevant literature, writing software, and running experiments—and must include either a substantial amount of programming or a strong theoretical mathematical treaty of issues related to multicore programming.
-    \item The final report is a short paper (6-10 pages) describing the results of the activity and should reflect the substantial amount of work that you put into the activity.
-The paper should make clear what is the problem, the state-of-the art in this research area, a discussion of related work, a clear definition of the contributions of your paper, a detailed discussion of your algorithms and approach and their application, and performed experimental results.
-The report will be graded for writing quality as well as for technical content, so attention should be paid to organization, grammar, spelling, and scholarly style.
-All references used, both text and code, must be properly cited in the bibliography.
-Your are required to use LaTeX (\url{http://en.wikipedia.org/wiki/LaTeX} (Links to an external site.)) to prepare your report.
-    \item Source code: typically in the course of your research you will develop source code for your algorithms and experimental results.
-Please include with your mid-term and final reports the following:
-    \begin{enumerate}
-        \item The source code of your algorithms along with a brief README file
-        \item The source code of your experiments along with a brief README file
-        \item The source code of your paper (in a LaTeX format).
-    \end{enumerate}
-\end{enumerate}
-\textit{Submission} \\
-All materials should be submitted online using Webcourses as a single zip or rar archive. \\
-\textit{Format} \\
-All papers should follow either the ACM or the IEEE Manuscript Format. \\
-A sample IEEE template and further formatting instructions are available here: \\
-\url{http://www.ieee.org/conferences_events/conferences/publishing/templates.html}
-
+\begin{thebibliography}{00}
+\bibitem{b1} G.
+Eason, B.
+Noble, and I.
+N.
+Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
+Trans.
+Roy.
+Soc.
+London, vol.
+A247, pp.
+529--551, April 1955.
+\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
+\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
+\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
+\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
+\bibitem{b6} Y.
+Yorozu, M.
+Hirano, K.
+Oka, and Y.
+Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
+J.
+Magn.
+Japan, vol.
+2, pp.
+740--741, August 1987 [Digests 9th Annual Conf.
+Magnetics Japan, p.
+301, 1982].
+\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\end{thebibliography}
 \end{document}

### 2022-03-25 paper.tex
Hash Table explanation

@@
 \usepackage{hyperref}
+\usepackage{amsmath}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \begin{document}
-\title{An Implementation of Councurrent Hash Tables in Java}
+\title{Implementation of Concurrent Hash Tables in Java}
 
@@
 \section{Research}
-Oh, that problem.
+\subsection{What are Hash Tables?}
+Hash tables are a data structure that utilizes an associative array, which is a structure that associates keys to specific values, in order to store data. The way
+in which a hash table uses these keys, is by entering them into a hash function. The hash function will then compute the array index in which
+the value associated with the key will be stored. The hash function is naturally the most important portion of the hash table's construction.
+The most commonly used hashing methods uses some variation of modular hashing. Modular hashing works as follows: assuming there is an array of size
+$M$, where $M$ is prime, then for any positive integer key $k$, compute $k \mod{M}$.
+
+Modular hashing is exapandable to accepting more than just integer as keys. A common data type used as a key for hash tables would be strings.
+In order to accomplish this, one would simply have to treat the string as though it were a large integer. This concept of converting data types into
+integers is commonly used as it is much simpler than creating a custom hashing function. A well made hash function will have three main requirements:
+deterministic - equivalent keys will always produce the same hash value, computationally efficient, and allows for keys to be uniformly distributed.
+Deterministic is only one of those requirements that is absolutely necessary for any hash function, as it what allows users to searh hash tables
+for a specific key-value pairing [1].
+
+There is of course, one issue with the hash function: since the hash table is not an array of infinite size, there is guaranteed to be instances where
+two distinct keys map to the same index when using the hash function. In order to resolve this, several techniques have been developed: linear probing,
+quadratic probing, double hashing, separate chaining or buckets. The basic concept for linear probing is that if the current key-value attempting to be
+entered into the hash table collides at position $i$, then attempt to place it at $i+1$. If it collides again, keep increasing the index by $1$ until
+it comes across an index in the hash table array that has not been used yet, ending at $i+n$ where $n$ is the number of collisions. Quadratic probing is
+similar to linear probing, except instead of ending up at $i+n$, the value will be stored at index $i+n^2$, where $n$ is the number of collisions.
+Double hashing, also known as rehashing, involves hashing the key a second time, with a different hash function. That result is taken, and then used as
+the step size for a process similar to the previous two methods, resulting in the value being place at index $i+jn$, where $n$ is the number of collisions,
+and $j$ is the result of the second hash function. These methods, known as open addresssing, work, but as the hash table begins to fill up, they result in the 
+add() operation taking more time to complete due to collisions. Separate chaining, on the other hand, works by having each index of the array be a linked list. 
+Whenever a value maps to thatindex, it will be added into list, which avoids the issue that open addressing presents. Buckets is a similar method to 
+separate chaining, using an array instead of a linked list. While separate chaining and buckets fix the issues with open addressing, they present more
+of their own, namely that there exist the possibility that a large majority of keys will map to the same index, which leads to a large amount of time
+being used searching the array or list at that index for the specified key's value [2].
 
@@
 \begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
+\bibitem{b1} R. Sedgewick and K. Wayne. "Hash Tables" princeton.edu \\https://algs4.cs.princeton.edu/34hash/ (accessed Mar. 13, 2022)
 \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.

### 2022-03-25 paper.tex
Manually Merging test-latex

@@
 
+\subsection{What is Concurrency?}
+Now given what a hash table is, this paper will look at what exactly it means for a computer program to be concurrent. Concurreny means that multiple computations
+will be occuring at the same time. This has become increasing more important, with the clock speed of processors no longer increasing, it results in 
+manufacturers adding more cores to chips, meaning splitting up computation among those chips is the only way to make it faster. Concurrent programming
+has two common models: shared memory models and message passing models. This paper will only look at shared memory models as the implementation for
+a concurrent hash table will utilize a shared memory model. A shared memory model works by having concurrent processors read and write objects into the
+memory that they share [3].
+
 \section{Related work}
@@
 
+\section{Appendix}
+
 \begin{thebibliography}{00}
-\bibitem{b1} R. Sedgewick and K. Wayne. "Hash Tables" princeton.edu \\https://algs4.cs.princeton.edu/34hash/ (accessed Mar. 13, 2022)
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\bibitem{b1} R. Sedgewick and K. Wayne. "Hash Tables." princeton.edu. \\https://algs4.cs.princeton.edu/34hash/ (accessed Mar. 13, 2022).
+\bibitem{b2} S. Mitra. "Hashing." utexas.edu. https://www.cs.utexas.edu/~mitra/\\csSpring2017/cs313/lectures/hash.html (accessed Mar. 13, 2022).
+\bibitem{b3} S. Amarasinghe, A. Chlipala, S. Devadas, M. Ernst, M. Goldman, J. Guttag, D. Jackson, R. Miller, M. Rinard, and A. Solar-Lezama. "Reading 17: Concurrency"
+mit.edu. https://web.mit.edu/6.005/\\www/fa14/classes/17-concurrency/ (accessed Mar. 14, 2022).
 \end{thebibliography}

### 2022-03-25 main/Experiment.java
Start of Experimentation

@@ -1,0 +1,32 @@
+import java.text.ParseException;
+import java.util.concurrent.TimeUnit;
+import java.util.concurrent.locks.*;
+
+public class Experiment {
+    public static void main(String[] args) throws ParseException, InterruptedException {
+        long startTime = System.nanoTime();
+        HashTable standard_unthreaded = new HashTable();
+        int su_insertion_failures  = 0; // How many times the insertion of a key-value failed.
+        int su_removal_failures = 0; // How many times the removal of a key-value failed.
+
+        for (int i = 0; i < 809; i++) {
+            standard_unthreaded.put(i);
+        }
+        for (int i = 0; i < 809; i++) {
+            if (!standard_unthreaded.search(i)) su_insertion_failures++;
+        }
+        for (int i = 0; i < 809; i++) {
+            standard_unthreaded.remove(i);
+        }
+        for (int i = 0; i < 809; i++) {
+            if (standard_unthreaded.search(i)) su_removal_failures++;
+        }
+        long endTime = System.nanoTime();
+        long duration = (endTime - startTime);
+
+        System.out.println(String.format("Unthreaded standard hashtable took %d nanoseconds to perform 809 inserts, 1618 searchs, and 809 deletes", duration));
+        System.out.println(String.format("Insertion failed %d times, and deletion failed %d times.\n", su_insertion_failures, su_removal_failures));
+
+        startTime = System.nanoTime();
+    }
+}

### 2022-03-25 main/LockingHashTable.java

@@ -1,11 +1,11 @@
-public class HashTable {
+public class LockingHashTable {
     public static final int ARR_SIZE = 809; // The maximum capacity of a HashTable
     private int[] table = new int[ARR_SIZE]; // Where values are stored
     private boolean[] nulls = new boolean[ARR_SIZE]; // If a value is null, it is true at the same index here
     private boolean[] cleans = new boolean[ARR_SIZE]; // Where clean indices are stored
     private int capacity;
 
-    public HashTable() {
+    public LockingHashTable() {
         // Generate arrays, each with default values. Set capacity to 0
         for (int i = 0; i < ARR_SIZE; i++) {
             table[i] = 0;
@@ -16,7 +16,7 @@ public class HashTable {
     }
 
     // Add an item to the hash table
-    public void put(int item) {
+    public synchronized void put(int item) {
         // If capacity if full, do not add
         if (capacity >= ARR_SIZE)
             return;
@@ -48,7 +48,7 @@ public class HashTable {
     }
 
     // Search the hash table for a specific item, return true if found, else false
-    public boolean search(int item) {
+    public synchronized boolean search(int item) {
         int key = hash(item);
         int i = 0;
         int new_key = (key + (i * i)) % ARR_SIZE;
@@ -68,7 +68,7 @@ public class HashTable {
     }
 
     // Removes an item from the hash table
-    public void remove(int item) {
+    public synchronized void remove(int item) {
         int key = hash(item);
         int i = 0;
         int new_key = (key + (i * i)) % ARR_SIZE;
@@ -127,5 +127,4 @@ public class HashTable {
 
         return s;
     }
-
 }

### 2022-03-25 main/Experiment.java
removed unused imports

@@ -1,6 +1,4 @@
 import java.text.ParseException;
-import java.util.concurrent.TimeUnit;
-import java.util.concurrent.locks.*;
 
 public class Experiment {
     public static void main(String[] args) throws ParseException, InterruptedException {

### 2022-03-25 paper.tex
Finished Appendix

@@
 \begin{abstract}
-Abstract goes here.
+
 \end{abstract}
@@
 
-\section{Related work}
-I've seen a similar problem before
-
-\section{Contributions}
-I've fixed this problem once
-
 \section{Algorithms}
@@
 \section{Experimental Results}
-Well, it worked on my machine.
+We begin by testing how the standard hash table will performed when run sequentially, and then when multithreaded.
 
 \section{Appendix}
+\subsection{Challenges}
+One of the major challenges the group came across was learning how to do concurrent programming within Java, as most of the group member's experience
+lay in C++.
+\subsection{Completed Tasks}
+\begin{itemize}
+    \item Creating a standard hash table.
+    \item Creating a hash table with locks.
+\end{itemize}
+\subsection{Remaining Tasks}
+\begin{itemize}
+    \item Creating a lockfree hash table.
+    \item More experimental testing.
+\end{itemize}
 

### 2022-03-25 paper.tex
Abstract added

@@
 \begin{abstract}
-
+    Concurrent programming is using multiple processes at the same time, often reducing total time waiting for operations to complete.
+A concurrent hash table is a hash table that utilizes concurrent programming in order to reduce the time it takes to do multiple operations, such as the add(), search(), and remove() methods hash tables utilize.
+It does this by doing different operations at the same time instead of one after another, waiting for the previous to finish before the next can even start.
+This report is about attempting to create our own concurrent hash table in the Java language instead of using Java’s own HashTable class.
+We will be achieving this by using function keywords like “synchronized” and manually locking and unlocking, experimenting will both options.
+We will show evidence of reduced time by running the same commands on an unthreaded custom HashTable and on a custom concurrent HashTable.
 \end{abstract}

### 2022-03-26 paper.tex
spelling and algorithm description

@@
 
-Modular hashing is exapandable to accepting more than just integer as keys. A common data type used as a key for hash tables would be strings.
+Modular hashing is expandable to accepting more than just integer as keys. A common data type used as a key for hash tables would be strings.
 In order to accomplish this, one would simply have to treat the string as though it were a large integer. This concept of converting data types into
@@
 deterministic - equivalent keys will always produce the same hash value, computationally efficient, and allows for keys to be uniformly distributed.
-Deterministic is only one of those requirements that is absolutely necessary for any hash function, as it what allows users to searh hash tables
+Deterministic is only one of those requirements that is absolutely necessary for any hash function, as it what allows users to search hash tables
 for a specific key-value pairing [1].
@@
 the step size for a process similar to the previous two methods, resulting in the value being place at index $i+jn$, where $n$ is the number of collisions,
-and $j$ is the result of the second hash function. These methods, known as open addresssing, work, but as the hash table begins to fill up, they result in the 
+and $j$ is the result of the second hash function. These methods, known as open addressing, work, but as the hash table begins to fill up, they result in the 
 add() operation taking more time to complete due to collisions. Separate chaining, on the other hand, works by having each index of the array be a linked list. 
-Whenever a value maps to thatindex, it will be added into list, which avoids the issue that open addressing presents. Buckets is a similar method to 
+Whenever a value maps to that index, it will be added into list, which avoids the issue that open addressing presents. Buckets is a similar method to 
 separate chaining, using an array instead of a linked list. While separate chaining and buckets fix the issues with open addressing, they present more
@@
 \subsection{What is Concurrency?}
-Now given what a hash table is, this paper will look at what exactly it means for a computer program to be concurrent. Concurreny means that multiple computations
-will be occuring at the same time. This has become increasing more important, with the clock speed of processors no longer increasing, it results in 
+Now given what a hash table is, this paper will look at what exactly it means for a computer program to be concurrent. Concurrency means that multiple computations
+will be occurring at the same time. This has become increasing more important, with the clock speed of processors no longer increasing, it results in 
 manufacturers adding more cores to chips, meaning splitting up computation among those chips is the only way to make it faster. Concurrent programming
@@
 \section{Algorithms}
-And this is how I did it!
-\begin{lstlisting}[language={Java},caption=Testing surce code in \LaTeX{}.,breaklines=true,frame=single]
-class HashTable<T> {
-    bool isConcurrent;
-    public static void main(String[] args) {
-        isConcurrent = true;
-    }
+% And this is how I did it!
+% \begin{lstlisting}[language={Java},caption=Testing surce code in \LaTeX{}.,breaklines=true,frame=single]
+% class HashTable<T> {
+%     bool isConcurrent;
+%     public static void main(String[] args) {
+%         isConcurrent = true;
+%     }
 
-    public HashTable<T>(bool isConcurrent) {
-        this.isConcurrent = isConcurrent;
-    }
-}
-\end{lstlisting}
+%     public HashTable<T>(bool isConcurrent) {
+%         this.isConcurrent = isConcurrent;
+%     }
+% }
+% \end{lstlisting}
+There are many ways to create single-threaded hash tables, so there are likewise many ways to create concurrent hash tables. If we ignore the need for resizing and
+the choice of hashing algorithm, we still need to decide on an approach for handling contention. We will attempt to employ methods such as course-grain locking, fine-grain locking,
+lazy, optimistic, and lock-free design.
 
@@
 \begin{itemize}
-    \item Creating a lockfree hash table.
+    \item Creating a lock-free hash table.
     \item More experimental testing.
@@
 \begin{thebibliography}{00}
-\bibitem{b1} R. Sedgewick and K. Wayne. "Hash Tables." princeton.edu. \\https://algs4.cs.princeton.edu/34hash/ (accessed Mar. 13, 2022).
-\bibitem{b2} S. Mitra. "Hashing." utexas.edu. https://www.cs.utexas.edu/~mitra/\\csSpring2017/cs313/lectures/hash.html (accessed Mar. 13, 2022).
-\bibitem{b3} S. Amarasinghe, A. Chlipala, S. Devadas, M. Ernst, M. Goldman, J. Guttag, D. Jackson, R. Miller, M. Rinard, and A. Solar-Lezama. "Reading 17: Concurrency"
+\bibitem{b1} R. Sedgewick and K. Wayne. ``Hash Tables." princeton.edu. \\https://algs4.cs.princeton.edu/34hash/ (accessed Mar. 13, 2022).
+\bibitem{b2} S. Mitra. ``Hashing." utexas.edu. https://www.cs.utexas.edu/~mitra/\\csSpring2017/cs313/lectures/hash.html (accessed Mar. 13, 2022).
+\bibitem{b3} S. Amarasinghe, A. Chlipala, S. Devadas, M. Ernst, M. Goldman, J. Guttag, D. Jackson, R. Miller, M. Rinard, and A. Solar-Lezama. ``Reading 17: Concurrency"
 mit.edu. https://web.mit.edu/6.005/\\www/fa14/classes/17-concurrency/ (accessed Mar. 14, 2022).

### 2022-03-29 main/Experiment.java
multithreaded experiment

@@ -2,29 +2,97 @@ import java.text.ParseException;
 
 public class Experiment {
     public static void main(String[] args) throws ParseException, InterruptedException {
-        long startTime = System.nanoTime();
-        HashTable standard_unthreaded = new HashTable();
-        int su_insertion_failures  = 0; // How many times the insertion of a key-value failed.
-        int su_removal_failures = 0; // How many times the removal of a key-value failed.
+        UnthreadedExperiment standard_u = new UnthreadedExperiment();
+        Thread single = new Thread(standard_u, "Single");
+        single.start();
+        single.join();
+        System.out.println("Finished unthreaded experiment on standard hash table.\n");
+        
+        ThreadedExperiment[] standard = new ThreadedExperiment[8];
+        Thread[] nonlocking = new Thread[8];
 
-        for (int i = 0; i < 809; i++) {
+        for (int i = 0; i < 8; i++) {
+            standard[i] = new ThreadedExperiment(i);
+            nonlocking[i] = new Thread(standard[i], Integer.toString(i));
+        }
+        
+        long startTime = System.nanoTime();
+        for (int i = 0; i < 8; i++) nonlocking[i].start();
+        for (int i = 0; i < 8; i++) nonlocking[i].join();
+        long endTime = System.nanoTime();
+        long duration = (endTime - startTime);
+
+        System.out.println(String.format("Threaded standard hashtable took %d milliseconds to perform 40024 inserts, 80048 searchs, and 40024 deletes", duration/1000000));
+        int insertion_failures = 0;
+        int removal_failures = 0;
+        for (int i = 0; i < 8; i++) {
+            insertion_failures += standard[i].insertion_failures;
+            removal_failures += standard[i].removal_failures;
+        }
+        System.out.println(String.format("Insertion failed %d times, and deletion failed %d times.", insertion_failures, removal_failures));
+        System.out.println("Finished threaded experiment on standard hash table.");
+    }
+}
+
+class UnthreadedExperiment implements Runnable {
+    HashTable standard_unthreaded;
+    int insertion_failures; // How many times the insertion of a key-value failed.
+    int removal_failures; // How many times the removal of a key-value failed.
+
+    UnthreadedExperiment() {
+        this.standard_unthreaded = new HashTable();
+        this.insertion_failures  = 0;
+        this.removal_failures = 0;
+    }
+
+    public void run() {
[…]

### 2022-03-29 main/HashTable.java

@@ -1,5 +1,5 @@
 public class HashTable {
-    public static final int ARR_SIZE = 809; // The maximum capacity of a HashTable
+    public static final int ARR_SIZE = 40024; // The maximum capacity of a HashTable
     private int[] table = new int[ARR_SIZE]; // Where values are stored
     private boolean[] nulls = new boolean[ARR_SIZE]; // If a value is null, it is true at the same index here
     private boolean[] cleans = new boolean[ARR_SIZE]; // Where clean indices are stored
@@ -106,7 +106,7 @@ public class HashTable {
 
     // Need to make a better hash function
     private int hash(int item) {
-        return item % 101;
+        return item % 5003;
     }
 
     // Prints the hash table in the same format as Java's own HashTable

### 2022-04-28 main/Experiment.java
Update Experiment.java

@@ -4,36 +4,42 @@ public class Experiment {
     public static void main(String[] args) throws ParseException, InterruptedException {
         for (int n = 0; n < 10; n++) {
             System.out.println("\n////////////////////////////////////////////////////////////////////////////////\n");
+
+            // Unthreaded
             UnthreadedExperiment standard_u = new UnthreadedExperiment();
             Thread single = new Thread(standard_u, "Single");
             single.start();
             single.join();
             System.out.println("Finished unthreaded experiment on standard hash table.\n");
-    /*
             ThreadedExperiment[] standard = new ThreadedExperiment[8];
             Thread[] nonlocking = new Thread[8];
 
+            // Nonlocking threaded
             for (int i = 0; i < 8; i++) {
                 standard[i] = new ThreadedExperiment(i);
                 nonlocking[i] = new Thread(standard[i], Integer.toString(i));
             }
-            
+
             long startTime = System.nanoTime();
-            for (int i = 0; i < 8; i++) nonlocking[i].start();
-            for (int i = 0; i < 8; i++) nonlocking[i].join();
+            for (int i = 0; i < 8; i++)
+                nonlocking[i].start();
+            for (int i = 0; i < 8; i++)
+                nonlocking[i].join();
             long endTime = System.nanoTime();
             long duration = (endTime - startTime);
 
-            System.out.println(String.format("Threaded standard hashtable took %d milliseconds to perform 64708 inserts, 129416 searches, and 64708 deletes", duration/1000000));
+            System.out.println(String.format(
+                    "Threaded standard hashtable took %d milliseconds to perform 64708 inserts, 129416 searches, and 64708 deletes",
+                    duration / 1000000));
             int insertion_failures = 0;
             int removal_failures = 0;
             for (int i = 0; i < 8; i++) {
                 insertion_failures += standard[i].insertion_failures;
                 removal_failures += standard[i].removal_failures;
             }
-            System.out.println(String.format("Insertion failed %d times, and deletion failed %d times.", insertion_failures, removal_failures));
+            System.out.println(String.format("Insertion failed %d times, and deletion failed %d times.",
+                    insertion_failures, removal_failures));
             System.out.println("Finished threaded experiment on standard hash table.\n");
-    */
 
             ThreadedExperiment2[] locking_experiment = new ThreadedExperiment2[8];
[…]

### 2022-04-28 main/Experiment.java
A lot of changes

@@ -1,36 +1,30 @@
 import java.text.ParseException;
+import java.util.Random;
 
 public class Experiment {
     public static void main(String[] args) throws ParseException, InterruptedException {
         for (int n = 0; n < 10; n++) {
-            System.out.println("\n////////////////////////////////////////////////////////////////////////////////\n");
-
-            // Unthreaded
-            UnthreadedExperiment standard_u = new UnthreadedExperiment();
-            Thread single = new Thread(standard_u, "Single");
-            single.start();
-            single.join();
-            System.out.println("Finished unthreaded experiment on standard hash table.\n");
-            ThreadedExperiment[] standard = new ThreadedExperiment[8];
+            ThreadedRandomExperiment[] standard = new ThreadedRandomExperiment[8];
             Thread[] nonlocking = new Thread[8];
-
+            System.out.println("\n////////////////////////////////////////////////////////////////////////////////\n");
             // Nonlocking threaded
             for (int i = 0; i < 8; i++) {
-                standard[i] = new ThreadedExperiment(i);
+                standard[i] = new ThreadedRandomExperiment(i);
                 nonlocking[i] = new Thread(standard[i], Integer.toString(i));
             }
 
             long startTime = System.nanoTime();
-            for (int i = 0; i < 8; i++)
+            for (int i = 0; i < 8; i++) {
                 nonlocking[i].start();
-            for (int i = 0; i < 8; i++)
+            }
+            for (int i = 0; i < 8; i++) {
                 nonlocking[i].join();
+            }
             long endTime = System.nanoTime();
             long duration = (endTime - startTime);
 
             System.out.println(String.format(
-                    "Threaded standard hashtable took %d milliseconds to perform 64708 inserts, 129416 searches, and 64708 deletes",
-                    duration / 1000000));
+                    "Randomized threaded standard hashtable took %d milliseconds", duration / 1000000));
             int insertion_failures = 0;
             int removal_failures = 0;
             for (int i = 0; i < 8; i++) {
@@ -40,36 +34,6 @@ public class Experiment {
             System.out.println(String.format("Insertion failed %d times, and deletion failed %d times.",
                     insertion_failures, removal_failures));
             System.out.println("Finished threaded experiment on standard hash table.\n");
-
-            ThreadedExperiment2[] locking_experiment = new ThreadedExperiment2[8];
-            Thread[] locking = new Thread[8];
-
-            for (int i = 0; i < 8; i++) {
[…]

### 2022-04-28 main/FineLockingHashTable.java

@@ -112,7 +112,7 @@ public class FineLockingHashTable implements HashTableInterface {
 
     // Need to make a better hash function
     private int hash(Integer item) {
-        return (item.hashCode()) % ARR_SIZE;
+        return Math.abs((item.hashCode()) % ARR_SIZE);
     }
 
     // Prints the hash table in the same format as Java's own HashTable

### 2022-04-28 main/HashTable.java

@@ -113,7 +113,7 @@ public class HashTable implements HashTableInterface {
 
     // Need to make a better hash function
     private int hash(Integer item) {
-        return (item.hashCode()) % ARR_SIZE;
+        return Math.abs((item.hashCode()) % ARR_SIZE);
     }
 
     // Prints the hash table in the same format as Java's own HashTable

### 2022-04-28 main/LockingHashTable.java

@@ -105,7 +105,7 @@ public class LockingHashTable implements HashTableInterface {
 
     // Need to make a better hash function
     private int hash(Integer item) {
-        return (item.hashCode()) % ARR_SIZE;
+        return Math.abs((item.hashCode()) % ARR_SIZE);
     }
 
     // Prints the hash table in the same format as Java's own HashTable

### 2022-04-28 main/Experiment.java
Better console logging and saving times

@@ -1,40 +1,98 @@
 import java.text.ParseException;
+import java.util.ArrayList;
 import java.util.Random;
+import java.util.concurrent.ConcurrentHashMap;
 
 public class Experiment {
+    static ArrayList<Long> unthreadedTimes = new ArrayList<Long>();
+    static ArrayList<Long> threadedTimes = new ArrayList<Long>();
+    static ArrayList<Long> lockingTimes = new ArrayList<Long>();
+
     public static void main(String[] args) throws ParseException, InterruptedException {
         for (int n = 0; n < 10; n++) {
-            ThreadedRandomExperiment[] standard = new ThreadedRandomExperiment[8];
+            System.out.println("//////////////////////////////////////////////////////");
+            // Unthreaded
+            UnthreadedExperiment standard_u = new UnthreadedExperiment();
+            Thread single = new Thread(standard_u, "Single");
+            single.start();
+            single.join();
+            // System.out.println("Finished unthreaded experiment on standard hash
+            // table.\n");
+            ThreadedExperiment[] standard = new ThreadedExperiment[8];
             Thread[] nonlocking = new Thread[8];
-            System.out.println("\n////////////////////////////////////////////////////////////////////////////////\n");
+
             // Nonlocking threaded
             for (int i = 0; i < 8; i++) {
-                standard[i] = new ThreadedRandomExperiment(i);
+                standard[i] = new ThreadedExperiment(i);
                 nonlocking[i] = new Thread(standard[i], Integer.toString(i));
             }
 
             long startTime = System.nanoTime();
-            for (int i = 0; i < 8; i++) {
+            for (int i = 0; i < 8; i++)
                 nonlocking[i].start();
-            }
-            for (int i = 0; i < 8; i++) {
+            for (int i = 0; i < 8; i++)
                 nonlocking[i].join();
-            }
             long endTime = System.nanoTime();
             long duration = (endTime - startTime);
 
-            System.out.println(String.format(
-                    "Randomized threaded standard hashtable took %d milliseconds", duration / 1000000));
+            /*
+             * System.out.println(String.format(
+             * "Threaded standard hashtable took %d milliseconds to perform 64708 inserts, 129416 searches, and 64708 deletes"
+             * ,
+             * duration / 1000000));
+             */
+            threadedTimes.add(duration / 1000000);
             int insertion_failures = 0;
             int removal_failures = 0;
             for (int i = 0; i < 8; i++) {
                 insertion_failures += standard[i].insertion_failures;
[…]

### 2022-04-28 main/Experiment.java
More experiments

@@ -1,3 +1,6 @@
+import java.io.BufferedWriter;
+import java.io.FileWriter;
+import java.io.IOException;
 import java.text.ParseException;
 import java.util.ArrayList;
 import java.util.Random;
@@ -9,7 +12,8 @@ public class Experiment {
     static ArrayList<Long> lockingTimes = new ArrayList<Long>();
 
     public static void main(String[] args) throws ParseException, InterruptedException {
-        for (int n = 0; n < 10; n++) {
+        int N = 100;
+        for (int n = 0; n < N; n++) {
             System.out.println("//////////////////////////////////////////////////////");
             // Unthreaded
             UnthreadedExperiment standard_u = new UnthreadedExperiment();
@@ -93,6 +97,28 @@ public class Experiment {
         System.out.println(unthreadedTimes);
         System.out.println(threadedTimes);
         System.out.println(lockingTimes);
+
+        try {
+            BufferedWriter out = new BufferedWriter(new FileWriter("experimentalTimes.txt"));
+            for (int i = 0; i < N - 1; i++) {
+                out.write(unthreadedTimes.get(i) + ",");
+            }
+            out.write(unthreadedTimes.get(N - 1) + "\n");
+
+            for (int i = 0; i < N - 1; i++) {
+                out.write(threadedTimes.get(i) + ",");
+            }
+            out.write(threadedTimes.get(N - 1) + "\n");
+
+            for (int i = 0; i < N - 1; i++) {
+                out.write(lockingTimes.get(i) + ",");
+            }
+            out.write(lockingTimes.get(N - 1) + "\n");
+
+            out.close();
+        } catch (IOException e) {
+            System.out.println("Failed to create file");
+        }
     }
 }
 

### 2022-04-28 main/Experiment.java
Experimentation evaluation

@@ -3,16 +3,16 @@ import java.io.FileWriter;
 import java.io.IOException;
 import java.text.ParseException;
 import java.util.ArrayList;
-import java.util.Random;
 import java.util.concurrent.ConcurrentHashMap;
 
 public class Experiment {
     static ArrayList<Long> unthreadedTimes = new ArrayList<Long>();
     static ArrayList<Long> threadedTimes = new ArrayList<Long>();
     static ArrayList<Long> lockingTimes = new ArrayList<Long>();
+    static ArrayList<Long> javaTimes = new ArrayList<Long>();
 
     public static void main(String[] args) throws ParseException, InterruptedException {
-        int N = 100;
+        int N = 3;
         for (int n = 0; n < N; n++) {
             System.out.println("//////////////////////////////////////////////////////");
             // Unthreaded
@@ -93,27 +93,52 @@ public class Experiment {
             // System.out.println("Finished threaded experiment on the locking hash
             // table.");
 
+            JavaConcurrent[] java_experiment = new JavaConcurrent[8];
+            Thread[] java_locking = new Thread[8];
+
+            for (int i = 0; i < 8; i++) {
+                java_experiment[i] = new JavaConcurrent(i);
+                java_locking[i] = new Thread(java_experiment[i], Integer.toString(i));
+            }
+
+            startTime = System.nanoTime();
+            for (int i = 0; i < 8; i++)
+                java_locking[i].start();
+            for (int i = 0; i < 8; i++)
+                java_locking[i].join();
+            endTime = System.nanoTime();
+            duration = (endTime - startTime);
+
+            /*
+             * System.out.println(String.format(
+             * "Threaded locking hashtable took %d milliseconds to perform 64708 inserts, 129416 searches, and 64708 deletes"
+             * ,
+             * duration / 1000000));
+             */
+            javaTimes.add(duration / 1000000);
+            insertion_failures = 0;
+            removal_failures = 0;
+            for (int i = 0; i < 8; i++) {
+                insertion_failures += java_experiment[i].insertion_failures;
+                removal_failures += java_experiment[i].removal_failures;
+            }
+            System.out
+                    .println(String.format("Java threaded insertion failed %d times, and deletion failed %d times.",
+                            insertion_failures, removal_failures));
+
         }
         System.out.println(unthreadedTimes);
         System.out.println(threadedTimes);
[…]

### 2022-04-28 main/evaluation.py

@@ -1,0 +1,11 @@
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+filename = "experimentalTimes.csv"
+
+df = pd.read_csv(filename)
+
+
+print(df["Unthreaded"])

### 2022-04-28 main/Experiment.java
More eval

@@ -12,7 +12,7 @@ public class Experiment {
     static ArrayList<Long> javaTimes = new ArrayList<Long>();
 
     public static void main(String[] args) throws ParseException, InterruptedException {
-        int N = 3;
+        int N = 10;
         for (int n = 0; n < N; n++) {
             System.out.println("//////////////////////////////////////////////////////");
             // Unthreaded

### 2022-04-28 main/evaluation.py

@@ -4,8 +4,19 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 filename = "experimentalTimes.csv"
-
 df = pd.read_csv(filename)
+linespace = range(1, len(df) + 1)
+names = df.columns
+rows, cols = 2, 2
+index = 0
 
 
-print(df["Unthreaded"])
+fig, axs = plt.subplots(ncols = cols, nrows = rows)
+for col in range(cols):
+	for row in range(rows):
+		sns.scatterplot(data = df, x = linespace, y = names[index], ax = axs[col][row]).set(title = names[index], ylabel = "Time (ms)")
+		axs[col][row].set_ylim(0, df[names[index]].max() * 1.1)
+		index += 1
+plt.subplots_adjust(left = 0.1, bottom = 0.1, right = 0.9, top = 0.9, wspace=.5, hspace=0.5)
+
+plt.show()

### 2022-04-28 paper.tex
Challenges

@@
 
+\subsection{Challenges of Concurrency}
+The largest challenge of creating any concurrent program is the concept of thread safety. For a data structure or algorithm to be consider threadsafe, it 
+must behave correctly when multiple threads are using it, without additional instructions from the function calling it. There are many strategies to utilize for
+creating a threadsafe data structure. One such strategy is known as \emph{confinement}, which avoids the issue of threads reading and writing to the same
+memory location because all the data is confined to a single thread, and the other threads do not have the ability to read or write to memory directly.
+Another strategy is \emph{immutability}, which means that the information which the threads share will be immutable, meaning they are unable to be altered by
+the threads. Other than those two, there is also the strategy of utilizing \emph{threadsafe data types}. Naturally utilizing data types that are already designed
+to be threadsafe helps when creating threadsafe algorithms and data structures because these data types are designed to not fail when multiple threads are being used
+[4]. There is one final strategy for desgining a threadsafe program, and that is to synchronized the threads so that they do not attempt to read or write to the same
+memory location at the same time. The most common way to implement thread synchronization in shared memory models is the use of locks. The way locks works is by taking
+the part of the program that accesses the data, and creating a "lock" around it. What this lock does is prevent multiple threads from being with in the "locked" region
+of the code. If a thread will attempt to enter the "locked" region, while another thread is already inside of it, the other thread will be a frozen 
+until the lock is removed and the thread will be able to access the region, reinitiating the lock again to prevent another thread from entering. Now that no other threads
+can enter the "locked" region, the thread inside of it is able to alter the memory location without having any other threads attempting to access it at the same time.
+
 \section{Algorithms}
-% And this is how I did it!
-% \begin{lstlisting}[language={Java},caption=Testing surce code in \LaTeX{}.,breaklines=true,frame=single]
-% class HashTable<T> {
-%     bool isConcurrent;
-%     public static void main(String[] args) {
-%         isConcurrent = true;
-%     }
-
-%     public HashTable<T>(bool isConcurrent) {
-%         this.isConcurrent = isConcurrent;
-%     }
-% }
-% \end{lstlisting}
 There are many ways to create single-threaded hash tables, so there are likewise many ways to create concurrent hash tables. If we ignore the need for resizing and
@@
 
-\section{Appendix}
-\subsection{Challenges}
-One of the major challenges the group came across was learning how to do concurrent programming within Java, as most of the group member's experience
-lay in C++.
-\subsection{Completed Tasks}
-\begin{itemize}
-    \item Creating a standard hash table.
-    \item Creating a hash table with locks.
-\end{itemize}
-\subsection{Remaining Tasks}
-\begin{itemize}
-    \item Creating a lock-free hash table.
-    \item More experimental testing.
-\end{itemize}
-
 \begin{thebibliography}{00}
@@
 mit.edu. https://web.mit.edu/6.005/\\www/fa14/classes/17-concurrency/ (accessed Mar. 14, 2022).
+\bibitem{b3} S. Amarasinghe, A. Chlipala, S. Devadas, M. Ernst, M. Goldman, J. Guttag, D. Jackson, R. Miller, M. Rinard, and A. Solar-Lezama. ``Reading 18: Thread Safety"
+mit.edu. http://web.mit.edu/6.005/\\www/fa14/classes/18-thread-safety/ (accessed Apr. 20, 2022).
+\bibitem{b3} S. Amarasinghe, A. Chlipala, S. Devadas, M. Ernst, M. Goldman, J. Guttag, D. Jackson, R. Miller, M. Rinard, and A. Solar-Lezama. ``Reading 20: Queues \& Locks"
+mit.edu. http://web.mit.edu/6.005/\\www/fa14/classes/20-queues-locks/(accessed Apr. 20, 2022).
 \end{thebibliography}

### 2022-04-28 main/Experiment.java
Better evaluation
Also ran the test 250 times and saved the times for each one.

@@ -12,7 +12,7 @@ public class Experiment {
     static ArrayList<Long> javaTimes = new ArrayList<Long>();
 
     public static void main(String[] args) throws ParseException, InterruptedException {
-        int N = 10;
+        int N = 250;
         for (int n = 0; n < N; n++) {
             System.out.println("//////////////////////////////////////////////////////");
             // Unthreaded

### 2022-04-28 main/evaluation.py

@@ -1,7 +1,14 @@
+import platform
 import numpy as np
 import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
+
+importedCPU = True
+try:
+	import cpuinfo
+except:
+	importedCPU = False
 
 filename = "experimentalTimes.csv"
 df = pd.read_csv(filename)
@@ -10,6 +17,16 @@ names = df.columns
 rows, cols = 2, 2
 index = 0
 
+if (importedCPU):
+	print(cpuinfo.cpu.info[0]["ProcessorNameString"])
+else:
+	print("Could not record CPU information")
+	
+
+print("Trials: " + str(len(df)))
+for name in names:
+	print("{0}:\n\tAverage: {1} ms\n\tHigh: {2} ms\n\tLow: {3} ms\n\tStandard deviation: {4} ms".
+		format(name, round(df[name].mean()), df[name].max(), df[name].min(), round(df[name].std(), 1)))
 
 fig, axs = plt.subplots(ncols = cols, nrows = rows)
 for col in range(cols):
@@ -17,6 +34,6 @@ for col in range(cols):
 		sns.scatterplot(data = df, x = linespace, y = names[index], ax = axs[col][row]).set(title = names[index], ylabel = "Time (ms)")
 		axs[col][row].set_ylim(0, df[names[index]].max() * 1.1)
 		index += 1
-plt.subplots_adjust(left = 0.1, bottom = 0.1, right = 0.9, top = 0.9, wspace=.5, hspace=0.5)
-
+fig.suptitle("Time Taken For {0} Tests".format(str(len(df))), fontsize = 15)
+plt.tight_layout()
 plt.show()

### 2022-04-29 paper.tex
Code for put()

@@
 \section{Algorithms}
-There are many ways to create single-threaded hash tables, so there are likewise many ways to create concurrent hash tables. If we ignore the need for resizing and
-the choice of hashing algorithm, we still need to decide on an approach for handling contention. We will attempt to employ methods such as course-grain locking, fine-grain locking,
-lazy, optimistic, and lock-free design.
+\subsection{Basic Hashtable}
+\begin{lstlisting}[language={Java},caption=Source code for the course-grain put() method.,captionpos=b,breaklines=true,frame=single]
+public synchronized boolean put(int item) {
+    // If capacity if full, do not add
+    if (capacity >= ARR_SIZE)
+        return false;
 
+    // Get key
+    int key = hash(item);
+
+    // Check if first try addition
+    if (table[key] == null) {
+        table[key] = item;
+        cleans[key] = false;
+    }
+    // Otherwise, quadratically probe for empty indices
+    else {
+        int i = 1;
+        int new_key = (key + i) % ARR_SIZE;
+        while (table[new_key] != null) {
+            i++;
+            new_key = (new_key + i) % ARR_SIZE;
+        }
+        table[new_key] = item;
+        cleans[new_key] = false;
+    }
+
+    // Increase capacity
+    capacity++;
+    return true;
+}
+\end{lstlisting}
 \section{Experimental Results}

### 2022-04-29 paper.tex
Added experimental data

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \section{Experimental Results}
-We begin by testing how the standard hash table will performed when run sequentially, and then when multithreaded.
+We began by testing how the standard hash table will performed when run sequentially, and then when multithreaded. 
+After those tests, we created a locking hash table and proceeded to run tests on that version. 
+After both of those tests, we deicded it would be a good idea to test our hash tables to Java's native concurrent hash map, the closest thing to a concurrent hash table Java has to offer natively.
+Our measurement for testing was timing how long each hash table takes to add a specific amount of numbers, make sure each one was added, remove all those numbers, and then make sure each was removed. 
+We recorded all the measurements into a CSV file and used a Python program to analyze the numbers, using Pandas to do the math and Matplotlib and Seaborn to visualize the data gathered. 
+We were able to gather a plentiful amount of information regarding each hash table and the time it takes to do these processes. 
+The data we gathered includes total time taken, the CPU used, the amount of tests, the time it took to do each test for each hash table, the average, low, and high times, and the standard deviation.
+All times recorded are in millisecconds.
+\begin{figure}[htbp]
+    \centerline{\includegraphics[width=90mm,scale=1]{250trials.png}}
+    \caption{Graphs of each test for each hash table. The x-axis is the test number, the y-axis is the time in ms the test took. }
+    \label{fig}
+\end{figure}
+
+\begin{table}[htbp]
+    \caption{Experimental Results}
+    \begin{center}
+    \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \textbf{Data}&\multicolumn{4}{|c|}{\textbf{Hash Table Version}} \\
+    \cline{2-5} 
+    \textbf{Calculation} & \textbf{\textit{Nonlocking}}& \textbf{\textit{Threaded}}& \textbf{\textit{Locked Threaded}}& \textbf{\textit{Java}} \\
+    \hline
+    \textbf{Average}& 8,135 ms& 130 ms&  5 ms& 272 ms\\
+    \hline
+    \textbf{High}& 10,978 ms& 145 ms&  66 ms& 1,499 ms\\
+    \hline
+    \textbf{Low}& 8164 ms& 118 ms&  4 ms& 247 ms\\
+    \hline
+    \textbf{Deviation}& 376.3 ms& 4.5 ms&  4.1 ms& 78.7 ms\\
+    \hline
+    \textbf{Total time}&\multicolumn{4}{|c|}{\textbf{2,180,629 ms}} \\
+    \hline
+    \multicolumn{5}{l}{Nonlocking and Threaded are the same class, but the latter was concurrently tested}
+    \end{tabular}
+    \label{tab1}
+    \end{center}
+\end{table}
 

### 2022-04-29 main/evaluation.py

@@ -1,33 +1,33 @@
-import platform
-import numpy as np
+# Modules
 import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
-
 importedCPU = True
 try:
 	import cpuinfo
 except:
 	importedCPU = False
 
+# Variables
 filename = "experimentalTimes.csv"
 df = pd.read_csv(filename)
 linespace = range(1, len(df) + 1)
 names = df.columns
 rows, cols = 2, 2
 index = 0
+totalTime = 0
 
+# Showing data
 if (importedCPU):
 	print(cpuinfo.cpu.info[0]["ProcessorNameString"])
 else:
 	print("Could not record CPU information")
-	
-
 print("Trials: " + str(len(df)))
 for name in names:
+	totalTime += df[name].sum()
 	print("{0}:\n\tAverage: {1} ms\n\tHigh: {2} ms\n\tLow: {3} ms\n\tStandard deviation: {4} ms".
 		format(name, round(df[name].mean()), df[name].max(), df[name].min(), round(df[name].std(), 1)))
-
+print("Total time to run {0} tests: {1} ms".format(str(len(df)), totalTime))
 fig, axs = plt.subplots(ncols = cols, nrows = rows)
 for col in range(cols):
 	for row in range(rows):

### 2022-04-29 paper.tex
Changes
Update paper.tex

@@
 \usepackage{graphicx}
+\usepackage{tabularx}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 \end{lstlisting}
+
+The code outlined in "Listing 1" is the put() method for the coarse-grain hash table created for this project. As it is a course grain hash table,
+it utilizes the \emph{synchronized} keyword to make a lock on this function, meaning that if one thread were to be in the middle of a call to this function
+then any thread that attempted to call it would have to wait until the original thread finished it's function call. Since this implementation uses a fixed size 
+hash table, the method will first check to see if the hash table has reached it's maximum capacity. If the hash table is full, then the method returns false,
+indicating that the attempt to put the item into the hash table was unsuccessful. In the case that the hash table still has unused memory, then the method
+will create a key for the item by hashing it, and checks to see if the memory location associated with that key has been used yet.
 \section{Experimental Results}
@@
     \begin{center}
-    \begin{tabular}{|c|c|c|c|c|}
+    \begin{tabularx}{\columnwidth}{|X|X|X|X|X|}
     \hline
@@
     \hline
-    \multicolumn{5}{l}{Nonlocking and Threaded are the same class, but the latter was concurrently tested}
-    \end{tabular}
+    \multicolumn{5}{l}{Threaded is Nonlocking but concurrently tested instead}
+    \end{tabularx}
     \label{tab1}

### 2022-04-29 paper.tex
Update paper.tex

@@
 
+Gathered from this data, we were able to come to some conclusions about the types of hash tables we created and the Java native concurrent hash map.
+The hash table that was created without concurrency in mind was slow, taking approximately 8 seconds per test. The deviation was relatively low at around 375 ms, 4\% of a difference from the average.
+The same hash table tested concurrently was much faster, able to do the same test but 60 times faster. The standard deviation, compared to the previous test, was significantly lower. 
+The test done after that one was the locking hash table, which performed significantly faster than even the concurrent non-locking one, completing the tests 26 times faster, 5 ms per test. 
+Contrary to previous thoughts, the Java hash map performed the second worst, averaging at 272 ms per test. \\
+These tests and theories gave us insight into some possible issues with our versions.
+The first of which, signified by the locking hash table tests, is that our testing methodology may not be perfect and have flaws that would make it seem good, but only in the case we gave it.
+Instead of thee same test everytime, we could randomize more numbers being added and removed, but adding that ability would slow the overall tests as we need to know which random numbers are being added and removed.
+That would slow down the tests themselves instead of changing the outcome of the tests. 
+While not a flaw, another phenominom that was noticed was the amount of time each hash table took to complete the first set of tests. 
+The unthreaded tests took, on average, 25\% longer to finish the first set of tests, eventually converging to about 8 seconds.
+The same can especially be seen in the locking and Java tests, taking approximately 16.5 and 5.5 times longer than the average.
+The speculated reason is the amount of time it takes for everything to be set up correctly the first time around. After that, the classes are already in memory and nothing new need to be set up.
+
 \begin{thebibliography}{00}

### 2022-04-29 paper.tex
run() explained

@@
 
+\begin{lstlisting}[language={Java},caption=Source code for the experiment used in each test,captionpos=b,breaklines=true,frame=single]
+    public void run() {
+        for (int i = (this.thread_num * 8089); i < (this.thread_num * 8089) + 8089; i++) {
+            table.put(i);
+        }
+        for (int i = (this.thread_num * 8089); i < (this.thread_num * 8089) + 8089; i++) {
+            if (!table.search(i))
+                insertion_failures++;
+        }
+        for (int i = (this.thread_num * 8089); i < (this.thread_num * 8089) + 8089; i++) {
+            table.remove(i);
+        }
+        for (int i = (this.thread_num * 8089); i < (this.thread_num * 8089) + 8089; i++) {
+            if (table.search(i))
+                removal_failures++;
+        }
+    }
+\end{lstlisting}
+
+In order to test if the hash tables work correctly, the \verb|run()| function was created, one for each experiment, each exactly the same, except for the Java hash map.
+The Java hash map's \verb|run()| function used \verb|table.put(i, i)| and \verb|table.contains(i)| in place of \verb|table.put(i)| and \verb|table.search(i)|, respectively. 
+In each test, each thread (for the threadded tests) was given a number, from 0 to the number of threads used in the test. The thread will then proceed to \verb|put()| in a series of numbers.
+After the series of numbers is added the the table, the thread will then, using the same series of numbers, search the table for those numbers.
+If the table failed to find a number, the insertion failures would increase, signified by the \verb|insertion_failures++;|.
+After that, the thread will then remove the same series of numbers from the table.
+Similar to the previous search, the thread will then search for the same series of numbers. However, if the number was found, the removal failed and the removal failure count would increment.
+This is signifed by the \verb|removal_failures++;| when \verb|table.search()| was successful.
+The variables \verb|insertional_failures| and \verb|removal_failures| are both public and static, so each thread and the overall experiment have access to it. 
+The \verb|table| variable, for each experiment, was different per hash table testing. For nonthreaded and locked, \verb|table = new HashTable()|.
+For Java hash maps, we set \verb|table| to \verb|ConcurrentHashMap<Integer, Integer>()| and for locked, we set it to \verb|LockingHashTable()|
+
 \begin{thebibliography}{00}

### 2022-04-29 paper.tex
Update paper.tex

@@
 \section{Algorithms}
-\subsection{Basic Hashtable}
+\subsection{Coarse-Grain Locking Hashtable}
 \begin{lstlisting}[language={Java},caption=Source code for the course-grain put() method.,captionpos=b,breaklines=true,frame=single]
@@
     }
-    // Otherwise, quadratically probe for empty indices
+    // Otherwise, linearly probe for empty indices
     else {
@@
 
-The code outlined in "Listing 1" is the put() method for the coarse-grain hash table created for this project. As it is a course grain hash table,
-it utilizes the \emph{synchronized} keyword to make a lock on this function, meaning that if one thread were to be in the middle of a call to this function
-then any thread that attempted to call it would have to wait until the original thread finished it's function call. Since this implementation uses a fixed size 
-hash table, the method will first check to see if the hash table has reached it's maximum capacity. If the hash table is full, then the method returns false,
+The code outlined in "Listing 1" is the \verb|put()| method for the coarse-grain hash table created for this project. As it is a course grain hash table,
+it utilizes the \verb|synchronized| keyword to make a lock on this method, meaning that if one thread were to be in the middle of a call to this method
+then any thread that attempted to call it would have to wait until the original thread finished it's method call. Since this implementation uses a fixed size hash table, 
+the method will first check to see if the hash table has reached it's maximum capacity. If the hash table is full, then the method returns false,
 indicating that the attempt to put the item into the hash table was unsuccessful. In the case that the hash table still has unused memory, then the method
-will create a key for the item by hashing it, and checks to see if the memory location associated with that key has been used yet.
+will create a key for the item by hashing it, and checks to see if the memory location associated with that key has been used yet. If no other key is associated
+with that memory location, then we add the item to the array at that location, and change the clean array so the hash table will know that data was stored 
+at this memory location for the linear probing in the \verb|remove()| method. If that memory location already has data associated with it, then the method will
+begin linear probing. The way linear probing works is by simply incrementing the key value by one, and checking that memory location. If the memory location
+is empty, then the method will store the item at that key value, otherwise, the method will continue to increment and check until it finds a memory location
+that is unused. After the value has been put in to the hash table, the hash table's capacity variable is increased to signify how many items are in it, and 
+the method call will finally return true to indicate that the value has been added successfully.
+
+\begin{lstlisting}[language={Java},caption=Source code for the course-grain search() method.,captionpos=b,breaklines=true,frame=single]
+public synchronized boolean search(int item) {
+    int key = hash(item);
+    int i = 0;
+    int new_key = (key + i) % ARR_SIZE;
+    // Quadratically probe for the item
+    // Dirty indices are considered to be a collision, the item may of been probed
+    // farther down
+    while (!cleans[new_key]) {
+        if (table[new_key] == null) {
+        } else if (item == table[new_key]) {
+            return true;
+        }
+        i++;
+        new_key = (new_key + i) % ARR_SIZE;
+        if (i >= ARR_SIZE)
+            break;
+    }
+
+    return false;
+}
+\end{lstlisting}
+
+The code outlined in "Listing 2" shows the \verb|search()| method for the coarse-grain hash table created for this project. This method's purpose is to accept 
+a value, and then check the locations in the hashtable where it could be stored for it. It does this by first acquiring the first key for that item with the
+\verb|hash()| method. After the method has the initial key associated with that value, it will check to see if that key is stored in the array at that location.
+It does this by creating a loop that checks to see the \verb|clean| array contains any data. What this accomplishes is a simple test, if there is no data in the \verb|clean| array
+then the item has never been put into the hash table, otherwise, there is a possibility that the value has been put in and is located here or further into
+the hash table due to linear probing. In the case that the locations checked in the \verb|clean| array are ever empty, the while loop is broken and the method returns
+false, indicating that the item has not been found. The while loop will continually check different locations inside the \verb|clean| array to see if they are equal to
+the item that is being searched for. If the item is found, the method returns true, otherwise the method will loop until every single memory location in the hash table has
+been checked and returns false.
+
+\begin{lstlisting}[language={Java},caption=Source code for the course-grain remove() method.,captionpos=b,breaklines=true,frame=single]
+public synchronized boolean remove(int item) {
+    int key = hash(item);
+    int i = 0;
+    int new_key = (key + i) % ARR_SIZE;
+
+    // If item not in hash table, no need to look for it again
+    if (!search(item)) {
+        return false;
+    }
+
+    // Quadratically probe for the item
+    // Dirty indices are considered to be a collision, the item may of been probed
+    // farther down
+    while (!cleans[new_key]) {
+        if (item == table[new_key]) {
+            table[new_key] = null;
+            break;
+        }
+        i++;
+        new_key = (new_key + i) % ARR_SIZE;
+    }
+
+    // Decrease capacity
+    capacity--;
+    return true;
+}
+\end{lstlisting}
+
 \section{Experimental Results}

### 2022-04-29 paper.tex
Typos
More typos fixed

@@
 \section{Experimental Results}
-We began by testing how the standard hash table will performed when run sequentially, and then when multithreaded. 
+We began by testing how the standard hash table will perform when run sequentially, and then when multi-threaded. 
 After those tests, we created a locking hash table and proceeded to run tests on that version. 
@@
 We recorded all the measurements into a CSV file and used a Python program to analyze the numbers, using Pandas to do the math and Matplotlib and Seaborn to visualize the data gathered.
+The file used to evaluate the data, "evaluation.py," can be found in the GitHub repository used for this project.
 We were able to gather a plentiful amount of information regarding each hash table and the time it takes to do these processes. 
@@
     \hline
-    \multicolumn{5}{l}{Threaded is Nonlocking but concurrently tested instead}
+    \multicolumn{5}{l}{CPU used for testing was a AMD Ryzen 5 3600X 6-Core Processor}
     \end{tabularx}

### 2022-04-29 paper.tex
Finished the Algorithm's Section

@@
     int new_key = (key + i) % ARR_SIZE;
-    // Quadratically probe for the item
+    // Linearly probe for the item
     // Dirty indices are considered to be a collision, the item may of been probed
@@
 
-    // Quadratically probe for the item
+    // Linearly probe for the item
     // Dirty indices are considered to be a collision, the item may of been probed
@@
 \end{lstlisting}
+
+"Listing 3" contains the code for coarse-grain locking hash table's \verb|remove()| method. The method is used to remove the value given to the method from the
+hash table. It begin's by calculating the initial key for that value by utilizing the hash table's \verb|hash()| method. After it calculates the key, the method
+makes a call to \verb|search()| to see if the value is in the hash table. If the value has not been added, then there is no way to remove it and the \verb|remove()|
+function returns false, due to the impossible nature of removing it. Now that the \verb|remove()| method is aware that the value it is trying to remove
+exists within the hash table, it needs to find and remove the value. It does this by utilizing the same exact loop from the \verb|search()| function, which 
+revolves around testing locations in the \verb|clean| to see if there has ever been data at this memory location in the hash table. The loop will continually check
+every single location that the data could have been placed into until it finds it, which the method is guaranteed to do as the call to \verb|search()| indicated
+that the value was in the array. Once the memory location with the value has been found, it is set to null in order to remove it from memory. After that, \verb|remove()|
+will decrease the \verb|capacity| variable, so that the hash table knows how many values are stored inside of it, and then returns true to indicate that the value
+was removed from the table successfully.
 

### 2022-04-29 paper.tex
Section For Fine-Grain Table

@@
 
+\subsection{Fine-Grain Locking Hashtable}
+
+\begin{lstlisting}[language={Java},caption=Source code for the fine-grain put() method.,captionpos=b,breaklines=true,frame=single]
+public boolean put(int item) {
+    // If capacity if full, do not add
+    if (capacity >= ARR_SIZE)
+        return false;
+
+    // Get key
+    int key = hash(item);
+
+    // Quadratically probe for empty indices
+    int i = 0;
+    int new_key = (key + (i * i)) % ARR_SIZE;
+    while (true) {
+        if (table[new_key] == null) {
+            locks[new_key].lock();
+            // Make sure still null
+            if (table[new_key] == null) {
+                table[new_key] = item;
+                cleans[new_key] = false;
+                locks[new_key].unlock();
+                break;
+            }
+            locks[new_key].unlock();
+        }
+
+        i++;
+        new_key = (new_key + (i * i)) % ARR_SIZE;
+    }
+
+    // Increase capacity
+    capacity++;
+    return true;
+}
+\end{lstlisting}
+
+\begin{lstlisting}[language={Java},caption=Source code for the fine-grain search() method.,captionpos=b,breaklines=true,frame=single]
+public boolean search(int item) {
+    int key = hash(item);
+    int i = 0;
+    int new_key = (key + (i * i)) % ARR_SIZE;
+    // Quadratically probe for the item
+    // Dirty indices are considered to be a collision, the item may of been probed
+    // farther down
+    while (!cleans[new_key]) {
+        if (item == table[new_key]) {
+            return true;
+        }
+        i++;
+        new_key = (new_key + (i * i)) % ARR_SIZE;
+        if (i >= ARR_SIZE)
+            break;
+    }
+
+    return false;
+}
+\end{lstlisting}
+
+\begin{lstlisting}[language={Java},caption=Source code for the fine-grain remove() method.,captionpos=b,breaklines=true,frame=single]
+public boolean remove(int item) {
+    int key = hash(item);
+    int i = 0;
+    int new_key = (key + (i * i)) % ARR_SIZE;
+
+    // If item not in hash table, no need to look for it again
+    if (!search(item)) {
+        return false;
+    }
+
+    // Quadratically probe for the item
+    // Dirty indices are considered to be a collision, the item may of been probed
+    // farther down
+    while (!cleans[new_key]) {
+        if (item == table[new_key]) {
+            locks[new_key].lock();
+            table[new_key] = null;
+            locks[new_key].unlock();
+            break;
+        }
+        i++;
+        new_key = (new_key + (i * i)) % ARR_SIZE;
+    }
+
+    // Decrease capacity
+    capacity--;
+    return true;
+}
+\end{lstlisting}
+
 \section{Experimental Results}

### 2022-04-29 main/Experiment.java
fixed threads not using same hashtable

@@ -193,13 +193,12 @@ class UnthreadedExperiment implements Runnable {
 }
 
 class ThreadedExperiment implements Runnable {
-    HashTable standard_unthreaded;
+    final HashTable standard_unthreaded = new HashTable();
     public int insertion_failures; // How many times the insertion of a key-value failed.
     public int removal_failures; // How many times the removal of a key-value failed.
     int thread_num;
 
     ThreadedExperiment(int thread_num) {
-        this.standard_unthreaded = new HashTable();
         this.insertion_failures = 0;
         this.removal_failures = 0;
         this.thread_num = thread_num;

### 2022-04-29 paper.tex
Update paper.tex

@@
 \end{lstlisting}
+For the fine-grain \verb|remove()| method, we start the same way as the other \verb|remove()| methods used in the other hash tables.
+Get the key and search the table for a value with that key. If it is not found, there is no need to try to remove the value, otherwise we just continue.
+If the table does have the value, we start to look for that value, and in this version, we quadratically probe to search for it.
+We continually check if the node is clean or not, dirty nodes mean there was never a value there and not worth checking onwards.
+When we encounter a clean node, we end.
+If we find the node we are searching for, we first start by locking it.
+After locking, we set the node to null and unlock it.
+After that, we break from the search and return true because we successfully removed the value.
+If we are still going through dirty nodes, we quadratically increase the index because we are quadratically probing. After all of that, we decrease the capacity of the table and return true.
 

### 2022-04-29 paper.tex
Finished Paper (hopefully)

@@
 
+The source code shown in "Listing 4" contains the \verb|put()| method for a fine-grain locking hash table. This method resembles the
+\verb|put()| method for the coarse-grain locking hash table, but differs in a few key ways. The first notable difference is the lack of 
+the \verb|syncrhonized| keyword in the method declaration. This means a call to this method would not block other threads from calling the
+method until the call finishes. Rather, this method implements the use of locks to block threads at the critical sections of the code, the
+critical section being the writing of the item into the hash table. What this code does is check to see if the memory location that the 
+data will be stored in is null, and then acquires the lock for that location. Once the lock is acquired, the method checks once again that
+memory location is null to ensure no other calls to the method wrote data in the time between checking and acquiring the lock. Once the data
+is added to the hash table, the lock will be relinquished and other thread will be able to continue with their method calls. The only other
+notable difference between the coarse-grain and fine-grain methods is that the fine-grain one utilizes quadratic probing.
+
 \begin{lstlisting}[language={Java},caption=Source code for the fine-grain search() method.,captionpos=b,breaklines=true,frame=single]
@@
 \end{lstlisting}
+
+The fine-grain \verb|search()| method is once again notably similar to its coarse-grain counterpart. The only difference between the two
+is difference in probing methods. The reasoning behind this is that since search only reads data from the array and does not write any,
+there is no reason for it to acquire a lock at any point in its runtime. The program has no reason to block access to the memory location
+in order to read it, and if data was being written to the memory location during the call to \verb|search()| then search would be blocked
+by \verb|put()| or \verb|remove| acquiring the lock to that memory location.
 

### 2022-04-29 paper.tex
paper

@@
 
+Follow up work:
+
+In the future, we might follow up this research by including other types of hash sets and performing more varied testing.
+It would be interesting to consider a lock-free tables, such as the lock-free hash set from The Art of Multiprocessor Programming.
+That table would likely have very different performance characteristics, given it's lock-free nature and unique overhead of maintaining a linked-list and buckets.
+We could also expand our testing to include different percentages of insertions, deletions, and reads. This might reveal other differences in performance.
+
 \begin{thebibliography}{00}
